Reduced diversity and increased specialization of Alaska salmon fishermen:

### 2017-03-05 salmon/paper/salmon_paper_v3.Rmd
paper and appendix changes

A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals. The modern era of Alaska’s large-scale hatchery program began in the late 1970s [@stopha2016], with substantial increases in hatchery production during the 1980s following low returns [@mahnken1998; @leber2008]. Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn2001a], though these programs may also have negative impacts on wild production [@wertheimer2005; @ruggerone2015]. The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn2001a; @hilborn1992a; @wertheimer2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber2008] or economic impacts [@boyce1993], rather than to variability experienced by individual fishermen. Understanding the role salmon have in buffering or stabilizing fishing revenues is critical given the responses of salmon to variable environments [@mueter2002; @schindler2010].

In addition to the changing species composition of fish available to be harvested, the benefits or costs of diversification are affected by a number of exogenous drivers, including competition from markets producing farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends]. Following price declines in the 1990s and early 2000s [@adfg2007], many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig. A8-A9). In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies). Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives or opportunities for diversification.

@@

We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC). Reported gross earnings (revenue) associated with each fish ticket were adjusted for inflation by converting all revenues too 2009 USD. Each fish ticket was associated with both a vessel and individual fisher, and because of our focus on individuals, we aggregated data by individuals to generate annual summaries of catch diversity and revenue. Because our modeling focus involves estimating time varying effects of diversification, we limited our analysis to salmon fisheries with species diversity > 1 (omitting fisheries such as the Bristol Bay gillnet fisheries that specialize on sockeye salmon). To better understand the influence of changing hatchery production, we also restricted fisheries to those that have been influenced by hatchery salmon production.

@@

To understanding long term change in variability of Alaska salmon fisheries, it is important to understand how effects change across scales, from individuals to fisheries to communities. For consistency with previous work, we examined the stability in the catch or landings in weight of salmon species [@hilborn2001a; @knapp2007], rather than revenue. We quantified the effects of changing hatchery production at three levels of data aggregation: (1) regions (combining fisheries), (2) between fisheries (defined by permits), and (3) within fisheries, for average fishers. We expect different results in examining the effects of fisheries versus individuals, because for a given fishery with no difference in landings over time, but changing participation, individuals would experience changing variability. We relied on previously published data as the response variable in comparisons across fisheries (Fig. 2), but the CFEC data for the remaining calculations. Hatchery release trends for some regions included in our data have been summarized previously (e.g. [@ward_evaluating] summarized the trends for Prince William Sound; [@stopha2016]; Fig. 3). Because hatchery contribution estimates to individual fisheries are calculated by region, rather than by fishery, we assumed that fisheries in the same region shared a common proportion of hatchery fish (e.g. seine and gillnet gears in Southeast Alaska).

To examine the effects of hatchery production on stability, we used a hierarchical variance function regression model. Variance function models (or heteroscedastic regression models, [@smyth1989]) are typically regression models that also allow covariates to be included (similar to performing sequential regressions on the mean and residuals [@western2009]). For a given level of aggregation (e.g. across fisheries) we modeled the mean catch of each species in each fishery (or region) as a AR(1) process, $$E[{ Y }_{ s,f,t }]={ Y }_{ s,f,t-1 }+{ b }_{ s,f }{ X }_{ s,f,t }+{ \phi  }_{ s,f }{ \varepsilon  }_{ s,f,t-1 }$$ where ${ Y }_{ s,f,t-1 }$ is the observed catch of species *s* in fishery *f* in the previous time step, ${ b }_{ s,f }$ is an estimated coefficient relating hatchery influence to the mean catch, and ${ X }_{ s,f,t }$ is the proportion of hatchery contribution to the total catch for species *s* in time *t* (0 to 1). The hatchery effect coefficients were modeled hierarchically, and assumed that ${ b }_{ s,f }\sim N\left( { u }_{ b },{ \sigma  }_{ b } \right)$, where ${ u }_{ b }$ and ${ \sigma  }_{ b }$ represent the location and scale hyperparameters. Finally, to reduce negative correlation in the residuals at lag-1, we included a moving average component with estimated parameter ${ \phi  }_{ s,f }$, where ${ \varepsilon  }_{ s,f,t-1 }$ represents the residual from the previous time step, ${ { \varepsilon  }_{ s,f,t-1 }=Y }_{ s,f,t-1 }-E\left[ { Y }_{ s,f,t-1 } \right]$.

To incorporate potential hatchery effects in the variability of the system, the second component of our model included a hierarchical component on the process variance terms, $$log\left( { \sigma  }_{ s,f,t } \right) ={ g  }_{ 0,s,f }+{ g  }_{ 1,s,f }{ X }_{ s,f,t }$$ where ${ g  }_{ 0,s,f }$ represents an estimated intercept unique to each fishery and species, and ${ g  }_{ 1,s,f }$ represents a fishery and species specific slope. We included the intercepts { g  }_{ 0,s,f } as fixed effects (representing differences by species) and again modeled the hatchery effect as hierarchical, ${ g }_{ 1,s,f }\sim N\left( { u }_{ g },{ \sigma  }_{ g } \right)$. The observed data were log-transformed and linked to the predicted mean and standard deviation with a normal likelihood, $ln\left( { Y }_{ s,f,t } \right) \sim N\left( E\left[ ln\left( { Y }_{ s,f,t } \right)  \right] ,{ \sigma  }_{ s,f,t } \right)$.

We focued inference on the slopes of the fraction of hatchery contribution to the total catch, ${ g  }_{ 1,s }$. To allow for comparison across fisheries and scales of data, we converted these estimates to effect sizes, by calculating ${ \kappa  }_{ s }=ln\left( \frac { { g }_{ 0,s }+{ \left( 1/3 \right) \cdot g }_{ 1,s } }{ { g }_{ 0,s } }  \right)$. The log ratio of these effect sizes represent the approximate natural log of the percent change in the standard deviation resulting from a change in the hatchery contribution of zero to 1/3 of the catch. 95% credible intervals were generated by repeating this calculation for each MCMC draw, and taking the quantiles of the resulting distribution.

### Have effects of diversification changed over time for salmon fishers?

Because we were only interested in single permit fisheries, we applied a simplified version of the CFEC data filtering steps described in [@anderson_benefits]. Individual trip-level catch and revenue statistics were aggregated to annual data, and to clarify our comparison across fisheries we retained individual fishers who only participated in one fishery in a year (Table 1). Annual revenues were log-transformed to make the data approximately normally distributed; to avoid complications in modeling individual switching permits, we only included pairs of years for a given person where they did not change permits (e.g. didn’t change fisheries between two years). [@anderson_benefits] implemented a hierarchical Bayesian time series model that incorporated group level predictors and individual covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers. This model also included random intercepts by fishery and year, allowing all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS).

@@

g1 = ggplot(people[people$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(n), color = Permit)) + geom_line() + ylab("Log10 unique people") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

@@

weight = left_join(weight, people)
g2 = ggplot(weight[weight$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(w/n), color = Permit)) + geom_line() + ylab("Log10 tons per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

@@

g3 = ggplot(dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(meanrev), color = Permit)) + geom_line() + ylab("Log10 revenue per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

g4 = ggplot(dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, meandiv, color = Permit)) + geom_line() + ylab("Species diversity") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

@@

```{r fig2, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2014. Average participation (unique fishers) given by circle size, and effective species diversity is shown with color shading. [EW notes: pretty cool clustering by permits here - highlighted by Ole. The modeling in the rest of the paper is largely comparing the blue seine fisheries to the orange gillnets in PWS / SEAK]", warning=FALSE, message=FALSE}
dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")

@@

seak_hatch$region = "Southeast"
pws_hatch$region = "Prince William Sound"
kodiak_hatch$region = "Kodiak"

@@

# bring in total harvest data (hatchery + wild)
hatch = read.csv("../../salmon/data-generated/commonPropertyCommercialCatch.csv")
hatch = hatch[,c("Year","Chinook","Sockeye","Coho","Pink","Chum","Region")]
hatch$Region = as.character(hatch$Region)
hatch$Region[which(hatch$Region=="PWS")] = "Prince William Sound"
hatch = melt(hatch, id.vars = c("Year","Region"))
names(hatch)[which(names(hatch)=="variable")]="Species"

g2 = ggplot(hatch, aes(Year, value, color=Species)) + 
  geom_line() + facet_wrap(~Region) + 
  xlab("Year") + ylab("% of total catch") + 
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum")) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

@@

```{r fig4, echo=FALSE, fig.pos="placeHere", fig.cap="Effects of the fraction of hatchery fish on the year-to-year variability in landed weight at three scales: (1) Combining fisheries across regions, (2) Across fisheries, and (3) Average variability of individual fishers within fisheries. Effects of variability are presented in terms of effect sizes, or the natural log of the standard deviation of year-to-year catches in a scenario with hatchery fish constituting 1/3 of the catch to the standard deviation in a scenario with no hatchery fish in the catch. Dots represent mean estimates, and error bars represent +/- 1 standard error"}

@@

df$fishery[which(df$fishery=="Prince William Sound")] = "            PWS"
df$fishery[which(df$fishery=="Southeast")] = "            SEAK"
df$fishery[which(df$fishery=="Kodiak")] = "            Kodiak"

p1 = ggplot(df, aes(x=fishery, y=estimate, group=species, color=species)) +
  geom_point(position=position_dodge(0.5), alpha = 0.5, size=2) +
  ylab("Ln ratio (effect size)") +
  xlab("") +
  ggtitle("Regions") +
  geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
    position=position_dodge(0.5), alpha=0.5) +
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
  geom_hline(aes(yintercept=0), linetype = 2) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

df = readRDS(file=paste0("../../salmon/data-generated/","fishery","_df.rds"))
df$fishery[which(df$fishery=="S 01A")]="SEAK-seine" 
df$fishery[which(df$fishery=="S 01E")]="PWS-seine" 
df$fishery[which(df$fishery=="S 01K")]="Kodiak-seine" 
df$fishery[which(df$fishery=="S 01M")]="AK Pen.-seine" 
df$fishery[which(df$fishery=="S 03A")]="SEAK-gillnet" 
df$fishery[which(df$fishery=="S 03E")]="PWS-gillnet" 
cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")

p2 = ggplot(df[df$fishery!="AK Pen.-seine",], aes(x=fishery, y=estimate, group=species, color=species)) +
  geom_point(position=position_dodge(0.5), alpha = 0.5, size=2) +
  ylab("Ln ratio (effect size)") +
  xlab("") +
  ggtitle("Fisheries") +
  geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
    position=position_dodge(0.5), alpha=0.5) +
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
  geom_hline(aes(yintercept=0), linetype = 2) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

df = readRDS(file=paste0("../../salmon/data-generated/","individual","_df.rds"))
df$fishery[which(df$fishery=="S 01A")]="SEAK-seine" 
df$fishery[which(df$fishery=="S 01E")]="PWS-seine" 
df$fishery[which(df$fishery=="S 01K")]="Kodiak-seine" 
df$fishery[which(df$fishery=="S 01M")]="AK Pen.-seine" 
df$fishery[which(df$fishery=="S 03A")]="SEAK-gillnet" 
df$fishery[which(df$fishery=="S 03E")]="PWS-gillnet"

p3 = ggplot(df[df$fishery!="AK Pen.-seine",], aes(x=fishery, y=estimate, group=species, color=species)) +
  geom_point(position=position_dodge(0.5), alpha = 0.5, size=2) +
  ylab("Ln ratio (effect size)") +
  xlab("") +
  ggtitle("Within fisheries") +
  geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
    position=position_dodge(0.5), alpha=0.5) +
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
  geom_hline(aes(yintercept=0), linetype = 2) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

gridExtra::grid.arrange(p1, p2, p3, ncol=2)

@@

```{r fig5, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1 (in other words, the effect of switching from effective catch diversity = 1 to 2). This shows declining benefits of diversification for most fisheries, S01M.", echo=FALSE}
# make yearpermit df
diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
years = seq(min(diffdat$year), max(diffdat$year))
permits = levels(as.factor(diffdat$permit))
# note that we have to use factor() here instead of unique() because
# that's the same as the ordering in the estimation code
yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
yearpermit$permit = substr(yearpermit$yrpermit,6,9)

@@

ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
  geom_line() +
  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

>>>>

The benefits or costs of catch diversification for fishermen are affected by a number of exogenous drivers, including competition from markets producing farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends]. Following price declines in the 1990s and early 2000s [@adfg2007], many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig. A8-A9). In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies). Perhaps one of the largest changes affecting the availability or incentives for fishermen to diversify has been the increase in the production of hatchery fish. The modern era of Alaska’s large-scale hatchery program began in the late 1970s [@stopha2016; @ward_evaluating], with substantial increases in hatchery production during the 1980s following low returns [@mahnken1998; @leber2008]. Many of these hatchery programs increased production to stabilize the variability of salmon catches [@hilborn2001a], though these programs may also have negative impacts on wild production [@wertheimer2005; @ruggerone2015]. The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn2001a; @hilborn1992a; @wertheimer2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber2008] or economic impacts [@boyce1993], rather than to variability experienced by fisheries or individual fishermen. Understanding the role salmon have in buffering or stabilizing fishing revenues is critical given the responses of salmon to variable environments [@mueter2002; @schindler2010].

@@

We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985-2016 from the Commercial Fisheries Entry Commission (CFEC). Reported gross earnings (revenue) associated with each fish ticket were adjusted for inflation by converting all revenues too 2009 USD. Each fish ticket was associated with both a vessel and individual fisher, and because of our focus on individuals, we aggregated data by individuals to generate annual summaries of catch diversity and revenue. Because our modeling focus involves estimating time varying effects of diversification, we limited our analysis to salmon fisheries with species diversity > 1 (omitting fisheries such as the Bristol Bay gillnet fisheries that specialize on sockeye salmon). To better understand the influence of changing hatchery production, we also restricted fisheries to those that have been influenced by hatchery salmon production.

@@

To understanding long term change in variability of Alaska salmon fisheries, it is important to understand how hatchery production has affected the mean revenue and variability in revenue. For consistency with previous work, we examined the stability in the catch (weight of salmon species [@hilborn2001a; @knapp2007], rather than revenue. Within each region in our analysis (PWS, Southeast, Kodiak, Alaska Peninsula), we used the hatchery contribution to the Common Property Commercial Harvest as the predictor variable, and total harvest as the response. These quantities are reported by region ([@stopha2016]; Fig. 3), but not at the level of fisheries or individuals. Because reporting was not done separately for the Kodiak and Alaska Peninsula regions until after 1995, we grouped these areas into a single region.

To examine the effects of hatchery production on stability, we used a hierarchical variance function regression model. Variance function models (or heteroscedastic regression models, [@smyth1989]) are typically regression models that also allow covariates to be included (similar to performing sequential regressions on the mean and residuals [@western2009]). We modeled the mean catch of each species in each region as an AR(1) process, $$E[{ Y }_{ s,f,t }]={ Y }_{ s,f,t-1 }+{ b }_{ s,f }{ X }_{ s,f,t }$$ where ${ Y }_{ s,f,t-1 }$ is the observed catch of species *s* in region *f* in the previous time step, ${ b }_{ s,f }$ is an estimated coefficient relating hatchery influence to the mean catch, and ${ X }_{ s,f,t }$ is the change in proportion of hatchery contribution to the total catch for species *s* in between times *t* and *t-1*. We used the difference in hatchery contribution, rather than raw values because the response variable was also first differenced. The hatchery effect coefficients were modeled hierarchically, and assumed that ${ b }_{ s,f }\sim N\left( { u }_{ b },{ \sigma  }_{ b } \right)$, where ${ u }_{ b }$ and ${ \sigma  }_{ b }$ represent the location and scale hyperparameters.

To incorporate potential hatchery effects in the variability of the system, the second component of our model included a hierarchical component on the process variance terms, $$log\left( { \sigma  }_{ s,f,t } \right) ={ g  }_{ 0,s,f }+{ g  }_{ 1,s,f }{ X }_{ s,f,t }$$ where ${ g  }_{ 0,s,f }$ represents an estimated intercept unique to each region and species, and ${ g  }_{ 1,s,f }$ represents a region and species specific slope. The covariate ${ X }_{ s,f,t }$ is identical to that used in the mean model. We included the intercepts { g  }_{ 0,s,f } as fixed effects (representing differences by species) and again modeled the hatchery effect as hierarchical, ${ g }_{ 1,s,f }\sim N\left( { u }_{ g },{ \sigma  }_{ g } \right)$. The observed data were log-transformed and linked to the predicted mean and standard deviation with a normal likelihood, $ln\left( { Y }_{ s,f,t } \right) \sim N\left( E\left[ ln\left( { Y }_{ s,f,t } \right)  \right] ,{ \sigma  }_{ s,f,t } \right)$.

We focued inference on the slopes of the fraction of hatchery contribution to the total catch, ${ g  }_{ 1,s }$. To allow for comparison across regions and scales of data, we converted these estimates to effect sizes, by calculating ${ \kappa  }_{ s }=ln\left( \frac { { g }_{ 0,s }+{ \left( 1/3 \right) \cdot g }_{ 1,s } }{ { g }_{ 0,s } }  \right)$. The log ratio of these effect sizes represent the approximate natural log of the percent change in the standard deviation resulting from a change in the hatchery contribution of zero to 10% of the catch. 95% credible intervals were generated by repeating this calculation for each MCMC draw, and taking the quantiles of the resulting distribution.

### Have effects of diversification changed over time for individual salmon fishers?

Because we were only interested in single permit fisheries, we applied a simplified version of the CFEC data filtering steps described in [@anderson_benefits]. Individual trip-level catch and revenue data 1975-2016 were aggregated to annual data for each fisher, and to clarify our comparison across fisheries we retained individual fishers who only participated in one fishery in a year (Table 1). Annual revenues were log-transformed to make the data approximately normally distributed; to avoid complications in modeling individual switching permits, we only included pairs of years for a given person where they did not change permits (e.g. didn’t change fisheries between two years). [@anderson_benefits] implemented a hierarchical Bayesian time series model that incorporated group level predictors and individual covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers. This model also included random intercepts by fishery and year, allowing all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS).

@@

people2 = people[people$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
people2$Permit[which(people2$Permit=="S01K")] = "Kodiak-seine"
people2$Permit[which(people2$Permit=="S01M")] = "AKP-seine"
people2$Permit[which(people2$Permit=="S01E")] = "PWS-seine"
people2$Permit[which(people2$Permit=="S01A")] = "SEAK-seine"
people2$Permit[which(people2$Permit=="S03A")] = "SEAK-gillnet"
people2$Permit[which(people2$Permit=="S03E")] = "PWS-gillnet"
people2$Permit[which(people2$Permit=="S03T")] = "BB-gillnet"
people2$Permit[which(people2$Permit=="S05B")] = "State-troll"

g1 = ggplot(people2, aes(year, log10(n), color = Permit)) + geom_line() + ylab("Log10 unique people") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

@@

weight = left_join(weight, people)

weight2 = weight[weight$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
weight2$Permit[which(weight2$Permit=="S01K")] = "Kodiak-seine"
weight2$Permit[which(weight2$Permit=="S01M")] = "AKP-seine"
weight2$Permit[which(weight2$Permit=="S01E")] = "PWS-seine"
weight2$Permit[which(weight2$Permit=="S01A")] = "SEAK-seine"
weight2$Permit[which(weight2$Permit=="S03A")] = "SEAK-gillnet"
weight2$Permit[which(weight2$Permit=="S03E")] = "PWS-gillnet"
weight2$Permit[which(weight2$Permit=="S03T")] = "BB-gillnet"
weight2$Permit[which(weight2$Permit=="S05B")] = "State-troll"

g2 = ggplot(weight2, aes(year, log10(w/n), color = Permit)) + geom_line() + ylab("Log10 tons per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

@@

dat.annual2 = dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
dat.annual2$Permit[which(dat.annual2$Permit=="S01K")] = "Kodiak-seine"
dat.annual2$Permit[which(dat.annual2$Permit=="S01M")] = "AKP-seine"
dat.annual2$Permit[which(dat.annual2$Permit=="S01E")] = "PWS-seine"
dat.annual2$Permit[which(dat.annual2$Permit=="S01A")] = "SEAK-seine"
dat.annual2$Permit[which(dat.annual2$Permit=="S03A")] = "SEAK-gillnet"
dat.annual2$Permit[which(dat.annual2$Permit=="S03E")] = "PWS-gillnet"
dat.annual2$Permit[which(dat.annual2$Permit=="S03T")] = "BB-gillnet"
dat.annual2$Permit[which(dat.annual2$Permit=="S05B")] = "State-troll"

g3 = ggplot(dat.annual2, aes(year, log10(meanrev), color = Permit)) + geom_line() + ylab("Log10 revenue per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

g4 = ggplot(dat.annual2, aes(year, meandiv, color = Permit)) + geom_line() + ylab("Species diversity") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

@@

```{r fig2, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2016. Average participation (unique fishers) given by circle size, and effective species diversity is shown with color shading. [EW notes: pretty cool clustering by permits here - highlighted by Ole. The modeling in the rest of the paper is largely comparing the blue seine fisheries to the orange gillnets in PWS / SEAK]", warning=FALSE, message=FALSE}
dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")

@@

seak_hatch$region = "Southeast"
pws_hatch$region = "Prince William Sound"
kodiak_hatch$region = "Kodiak - Alaska Peninsula"

@@

# bring in total harvest data (hatchery + wild)
hatch = read.csv("../../salmon/data-generated/commonPropertyCommercialCatch.csv",
  stringsAsFactors = FALSE)

kodap1 = hatch[which(hatch$Region%in%c("Kodiak") & hatch$Year <= 1995),]
kodap2 = hatch[which(hatch$Region%in%c("AP", "Kodiak") & hatch$Year > 1995),]
kodap2 = dplyr::group_by(kodap2, Year) %>% 
  summarize(Chinook_harvest = sum(Chinook_harvest),
    Sockeye_harvest = sum(Sockeye_harvest),
    Coho_harvest = sum(Coho_harvest),
    Pink_harvest = sum(Pink_harvest),
    Chum_harvest = sum(Chum_harvest),
    Region = Region[1],
    Source = Source[1],
    ReportNumber = ReportNumber[1],
    Chinook_enhanced = sum(Chinook_enhanced),
    Sockeye_enhanced = sum(Sockeye_enhanced),
    Coho_enhanced = sum(Coho_enhanced),
    Pink_enhanced = sum(Pink_enhanced),
    Chum_enhanced = sum(Chum_enhanced))
kodap = rbind(as.data.frame(kodap1), as.data.frame(kodap2))
kodap$Region = as.character(kodap$Region)
kodap$Source = as.character(kodap$Source)
kodap$ReportNumber = as.character(kodap$ReportNumber)

hatch = rbind(hatch[which(hatch$Region %in% c("PWS","Southeast")),], kodap)

hatch$pChinook = hatch$Chinook_enhanced/hatch$Chinook_harvest
hatch$pSockeye = hatch$Sockeye_enhanced/hatch$Sockeye_harvest
hatch$pCoho = hatch$Coho_enhanced/hatch$Coho_harvest
hatch$pPink = hatch$Pink_enhanced/hatch$Pink_harvest
hatch$pChum = hatch$Chum_enhanced/hatch$Chum_harvest

hatch = hatch[,c("Year","pChinook","pSockeye","pCoho","pPink","pChum","Region")]
hatch = dplyr::rename(hatch, Chinook=pChinook,Sockeye=pSockeye, Coho=pCoho, Pink=pPink, Chum=pChum)
hatch$Region = as.character(hatch$Region)
hatch$Region[which(hatch$Region=="PWS")] = "Prince William Sound"
hatch$Region[which(hatch$Region=="Kodiak")] = "Kodiak - Alaska Peninsula"
hatch = melt(hatch, id.vars = c("Year","Region"))
names(hatch)[which(names(hatch)=="variable")]="Species"

g2 = ggplot(hatch, aes(Year, value, color=Species)) + 
  geom_line() + facet_wrap(~Region) + 
  xlab("Year") + ylab("Hatchery fraction of total catch") + 
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum")) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

@@

```{r fig4, echo=FALSE, fig.pos="placeHere", fig.cap="Effects of the fraction of hatchery fish on the year-to-year variability in landed weight at three scales: (1) Combining fisheries across regions, (2) Across fisheries, and (3) Average variability of individual fishers within fisheries. Effects of variability are presented in terms of effect sizes, or the natural log of the standard deviation of year-to-year catches in a scenario with hatchery fish constituting 10% of the catch to the standard deviation in a scenario with no hatchery fish in the catch. Dots represent mean estimates, and error bars represent +/- 1 standard error"}

@@

df$fishery[which(df$fishery=="Prince William Sound")] = "PWS"
df$fishery[which(df$fishery=="Southeast")] = "SEAK"
df$fishery[which(df$fishery=="Kodiak")] = "Kodiak"

df$fishery[which(df$fishery=="Prince William Sound")] = "PWS"
df$fishery[which(df$fishery=="Southeast")] = "SEAK"
df$fishery[which(df$fishery=="Kodiak")] = "Kodiak"

p1 = ggplot(df, aes(x=fishery, y=bestimate, group=species, color=species)) +
  geom_point(position=position_dodge(0.5), alpha = 0.5, size=3) +
  ylab("Effect size (mean)") +
  xlab("") +
  geom_errorbar(aes(ymin=bestimate-bsd, ymax=bestimate+bsd),
    position=position_dodge(0.5), alpha=0.5) +
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
  geom_hline(aes(yintercept=0), linetype = 2) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

p2 = ggplot(df, aes(x=fishery, y=estimate, group=species, color=species)) +
  geom_point(position=position_dodge(0.5), alpha = 0.5, size=3) +
  ylab("Effect size (sd)") +
  xlab("") +
  geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
    position=position_dodge(0.5), alpha=0.5) +
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
  geom_hline(aes(yintercept=0), linetype = 2) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

gridExtra::grid.arrange(p1, p2, ncol=1)
```

@@

```{r fig5, fig.pos="placeHere", fig.cap="Time varying effects of increasing the effective diversity of catch from 1 to 2 species (measured in ln dollars). Effective species diversity is calculated as 1/Simpson's Diversity. This shows declining benefits of diversification for most fisheries.", echo=FALSE}
# make yearpermit df
diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
years = seq(min(diffdat$year), max(diffdat$year))
permits = levels(as.factor(diffdat$permit))
# note that we have to use factor() here instead of unique() because
# that's the same as the ordering in the estimation code
yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
yearpermit$permit = substr(yearpermit$yrpermit,6,9)

@@

df$permits[which(df$permits=="S01A")] = "SEAK-seine"
df$permits[which(df$permits=="S01E")] = "PWS-seine"
df$permits[which(df$permits=="S01K")] = "Kodiak-seine"
df$permits[which(df$permits=="S01M")] = "AKP-seine"
df$permits[which(df$permits=="S03A")] = "SEAK-gillnet"
df$permits[which(df$permits=="S03E")] = "PWS-gillnet"
